Report on laboratory assignments 1 and 2 for EDA282, \\ Parallel Computer Organization and Design

### 2014-10-14 lab/report/include/lab1.tex

\subsection{Task 1}
\label{sec:lab11}
Three different microbenchmarks with disgustingly various communication patterns are used in the simulations and here follow a description of each benchmark.
\subsubsection*{Benchmark 1 - Producer consumer}
In this benchmark, one thread increments the values in the counter array, whereas the other threads read these values and accumulate the read values in a private variable. However, it is not ensured that the incrementing thread increments the values before the other threads reads, and thus can this happen in any order. The counter accesses is done in mutual exclusion using a lock, and before a new lap of increments \& reads there is a barrier.

\subsubsection*{Benchmark 2 - No sharing}
This benchmark have no sharing of the counter array. The counter array is divided up equally among the threads, giving each thread access to every nth counter element, where n is the number of threads. Each thread then increments it's assigned counters sequentially. A barrier is placed before a new incrementation lap of the assigned array elements begins.

@@

\subsection{Task 2}
\label{sec:lab12}
Here the CCPs MSI and MESI are compared using benchmarks 1 and 2. The difference between the CCPs is that MESI includes an exclusive (E) state, which is beneficial if data is read exclusively by one processor. If the data is then read by another processor, a move to the shared (S) state is made without cost. I.e. MESI can only produce equal or better results than MSI.

In the producer consumer benchmark, the reading threads transition the data element to the E state when accessing the array for the first time if the writing thread have not accessed it already. However, the whole counter array is shared and thus will not state E be of any gain. Therefore, MSI and MESI performs equally regardless of working set size, which is shown in \reffig{fig:lab1bars}.

@@

\subsection{Task 3}
\label{sec:lab13}
MESI is here compared with MESI-MG by executing benchmark 1 and 3. MESI-MG is a modified version of MESI where the S state is not utilized. If a cache issues a read request of a data element which exists in state E or M in a remote cache, the data is forwarded from the remote cache and then the remote cache invalidates it's copy. By this scheme, the S state is never used.

@@

In the all shared benchmark, every thread increments every data element during mutual exclusion which favors MESI-MG. Each data element is read and modified exclusively by one processor at a time, thus is the S state not beneficial. As MESI-MG do not utilize the S state, as described, it is more suitable than MESI in this benchmark. In Appendix [???] it is shown that the CCPs results are a like, except that MESI have to issue a significant number of invalidates, which is needed for the transition from state S to M. As shown in \reffig{fig:lab1bars}, MESI-MG's benefit is not visible for the working set size increasing the size of the LLC. That is because the cycle count is dominated by the LLC misses, as in the producer consumer benchmark.

>>>>

\subsection{Task 1}
\label{sec:lab11}
Three different microbenchmarks with disgustingly various communication patterns are used in the simulations and here follow a description of each.
\subsubsection*{Benchmark 1 - Producer consumer}
In this benchmark, one thread increments the values in the counter array, whereas the other threads read these values and accumulate the read values in a private variable. However, it is not ensured that the incrementing thread increments the values before the other threads reads, and thus can this happen in any order. A counter access is done in mutual exclusion using a lock, and before a new lap of increments \& reads there is a barrier.

\subsubsection*{Benchmark 2 - No sharing}
This benchmark have no sharing of the counter array. The counter array is divided up equally among the threads, giving each thread access to every nth counter element, where n is the number of threads. Each thread then increments it's assigned counters sequentially. A barrier is placed before a new incrementation lap of the assigned array elements begin.

@@

\subsection{Task 2}
\label{sec:lab12}
Here the CCPs MSI and MESI are compared using benchmark 1 and 2. The difference between the CCPs is that MESI includes an exclusive (E) state, which is beneficial if data is read exclusively by one processor. If the data is then read by another processor, a move to the shared (S) state is made without cost. I.e. MESI can only produce equal or better results than MSI.

\todo[inline]{change}
In the producer consumer benchmark, the reading threads transition the data element to the E state if the writing thread have not accessed it already when accessing the array for the first time or when fetching it from higher level caches. However, the whole counter array is shared and thus will not state E be of any gain. Therefore, MSI and MESI performs equally regardless of working set size, which is shown in \reffig{fig:lab1bars}.

@@

\subsection{Task 3}
\label{sec:lab13}
MESI is here compared with MESI-MG by executing benchmark 1 and 3. MESI-MG is a modified version of MESI where the S state is not utilized; if a cache issues a read request of a data element which exists in state E or M in a remote cache, the data is forwarded from the remote cache and then the remote cache invalidates it's copy. By this scheme, the S state is never used.

@@

In the all shared benchmark, every thread increments every data element in mutual exclusion which favors MESI-MG. Each data element is read and modified exclusively by one processor at a time, thus is the S state not beneficial. As MESI-MG do not utilize the S state, as described, it is more suitable than MESI in this benchmark. In Table \ref{tbl:lab1:task3} it is shown that the CCPs results are alike, except that MESI have to issue a significant number of invalidates, which is needed for the transition from state S to M. As shown in \reffig{fig:lab1bars}, MESI-MG's benefit is not visible for the working set size increasing the size of the LLC. That is because the cycle count is dominated by the LLC misses, as in the producer consumer benchmark.
